---
title: "Antarctica - ASV analysis "
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

# Initialize

## Libraries

```{r, echo=TRUE, message=FALSE, warning=FALSE}
   if(any(grepl("package:dvutils", search()))) detach("package:dvutils", unload=TRUE)
   library(dvutils)
   library(stringr)
   library(ggplot2)
   library(dplyr)
   library(tidyr)
   library(tibble)
   library(readr)
   library(maps)
   library(glue)
   library(readxl)
   library(rlang)
   
   library(phyloseq)

   library(knitr)
   library(rmdformats)

   # library(pr2database)
   # data("pr2")
```

## Markdown
```{r knitr_init, echo=FALSE, cache=FALSE}


## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               eval = TRUE,
               rows.print=50)
opts_knit$set(width=75)
```

## Read the colors

```{r define_colors}

 class_colors.df <- read_excel("class_colors.xlsx", sheet = "classes")
  
 class_colors <-structure(class_colors.df$color_name,.Names=class_colors.df$class) 
 
```


# Get the Antarctica ASVs


|set id |  description |
|---|---|
| 16 | Antar_2015_18S_V4 |
| 17 | Antar_2015_16S_plastid |
| 18 | Antar_2015_18S_V4_sorted |
	
* Only use asv for which supergroup_boot >= 90
* Only keep photosynthetic groups abd exclude dinoflagellates

```{r, message=FALSE}
  # rm(list = ls())
  sample_type <- c("18S filter", "16S plastid", "18S sort")
  taxon_selected <- c("Chlorophyta", "Cryptophyta", "Rhodophyta","Haptophyta", "Ochrophyta")
  
  asv <- list()
  ps <- list()
  long <- list()
# Export as specific data set as a phyloseq file
  
  i <- 16
  for (one_sample_type in sample_type) {
      asv[[one_sample_type]] <- metapr2_export_asv(dataset_id_selected = i, 
                                                   taxo_level = division, 
                                                   taxo_name = taxon_selected,
                                                   export_phyloseq = TRUE, 
                                                   export_long_xls = FALSE,
                                                   export_wide_xls = FALSE,
                                                   export_fasta = FALSE,
                                                   export_sample_xls = FALSE,
                                                   boot_min = 90, 
                                                   boot_level = supergroup_boot,
                                                   directory = "../dada2/") 

      ps[[one_sample_type]] <- asv[[one_sample_type]][["ps"]] 
      
  # Label samples with date 
      sample_data(ps[[one_sample_type]])$sample_label <- as.character(sample_data(ps[[one_sample_type]])$date)
      
  # Add TFF to sample name
      # sample_data(ps[[one_sample_type]])$sample_label <- str_c(sample_data(ps[[one_sample_type]])$metadata_code, 
      #                                                          str_replace_na(sample_data(ps[[one_sample_type]])$sample_concentration, ""), 
      #                                                          sep = "." )
      
      print(glue("Phyloseq - {one_sample_type}"))
      print(ps[[one_sample_type]] )
      cat("============================\n")
      
  i <- i + 1
  
  }
  
  
  sample_variables(ps[["18S filter"]])
  
  rm(asv)

```



# Get the different fraction separately

```{r, message=FALSE}


  fraction_filter <- c(0.2, 3, 20)
  fraction_sort <- c("pico", "nano")
  
  sample_type_filtered <- c(str_c(sample_type[1], fraction_filter, "um", sep=" "), 
                            str_c(sample_type[2], fraction_filter, "um", sep=" "),
                            str_c(sample_type[3], fraction_sort, sep=" "))
  
# Export as specific data set as a phyloseq file
  
  for (one_sample_type in sample_type[1:2]) {
    for (one_fraction in fraction_filter) {
      one_sample_type_filtered <- c(str_c(one_sample_type, one_fraction, "um", sep=" "))
      ps[[one_sample_type_filtered]] <- ps[[one_sample_type]] %>% 
        subset_samples(fraction_min == one_fraction) %>% 
        filter_taxa(function(x) sum(x) > 0 , TRUE)
      
      print(glue("Phyloseq - {one_sample_type_filtered}"))
      print(ps[[one_sample_type_filtered]] )
      cat("============================\n")
    }
  
  }
  
# Sorted samples
  
    for (one_sample_type in sample_type[3]) {
    for (one_fraction in fraction_sort) {
      one_sample_type_filtered <- c(str_c(one_sample_type, one_fraction, sep=" "))
      ps[[one_sample_type_filtered]] <- ps[[one_sample_type]] %>% 
        subset_samples(fraction_name == one_fraction) %>% 
        filter_taxa(function(x) sum(x) > 0 , TRUE)
      
      print(glue("Phyloseq - {one_sample_type_filtered}"))
      print(ps[[one_sample_type_filtered]] )
      cat("============================\n")
    }
  
    }
  
  sample_type_all <- c(sample_type, sample_type_filtered)
  

```
# Surface samples

* Only use Station 6 (not 14)
* Only surface considered (5 m)
* Do not consider TFF samples 
* Very important, must remove taxa that are not present in the filtered samples

```{r}
  sample_type_surface <- str_c(sample_type_all, " surface")
  
  for (one_sample_type in sample_type_all) {
      
      one_sample_type_surface = str_c(one_sample_type, " surface")

      ps[[one_sample_type_surface]] <- ps[[one_sample_type]] %>% 
        subset_samples(station_id=="6") %>% 
        subset_samples(depth_level=="surface") %>% 
        subset_samples(is.na(sample_concentration)) %>%
        filter_taxa(function(x) sum(x) > 0 , TRUE)
      
      # Rename the samples
      
      if (!str_detect(one_sample_type_surface, "sort")){
        sample_names(ps[[one_sample_type_surface]]) <- str_c(as.character(sample_data(ps[[one_sample_type_surface]])$date), 
                                               as.character(sample_data(ps[[one_sample_type_surface]])$depth), "m",
                                               as.character(sample_data(ps[[one_sample_type_surface]])$fraction_min),
                                               sep="_")
      } else {
        sample_names(ps[[one_sample_type_surface]]) <- str_c(as.character(sample_data(ps[[one_sample_type_surface]])$date), 
                                       as.character(sample_data(ps[[one_sample_type_surface]])$depth), "m",
                                       as.character(sample_data(ps[[one_sample_type_surface]])$fraction_name),
                                       sep="_")
      }
      
      
      print(glue("Phyloseq - {one_sample_type_surface}"))
      print(ps[[one_sample_type_surface]] )
      cat("============================\n")
      
  }

```
## Normalize and transform to long form

```{r, message=FALSE}

  for (one_sample_type in sample_type_surface) {
    
      cat(one_sample_type)
      cat("\n============================")  
      
      ps[[one_sample_type]] <- phyloseq_normalize_median(ps[[one_sample_type]])
      
      long[[one_sample_type]] <- phyloseq_transform_to_long(ps[[one_sample_type]])
      
      cat("\n") 
      
  }


```

## List of classes

```{r fig.height=6, fig.width=6}

  classes <- data.frame(division = NULL, class= NULL)
  
  for (one_sample_type in sample_type_surface[1:3]) {
      classes <- bind_rows(classes, select(long[[one_sample_type]], division, class))
  }

      classes <- classes %>% 
        distinct(division, class) %>% 
        arrange(division, class)
      
      # print(classes, n=50)
      knitr::kable(classes)

```


## Treemaps


```{r fig.height=6, fig.width=6}

  treemap_class <- list()
  treemap_genus <- list()
  
  for (one_sample_type in sample_type_surface) {
    
      treemap_class[[one_sample_type]] <- phyloseq_long_treemap(long[[one_sample_type]], 
                                                                     division, class, 
                                                                     str_c(one_sample_type, " - Class" ),
                                                                     colors = class_colors)
      treemap_genus[[one_sample_type]] <- phyloseq_long_treemap(long[[one_sample_type]], 
                                                                     class, genus, 
                                                                     str_c(one_sample_type, " - Genus" ))
 
  }

```

## ASVs bargraphs


```{r, fig.height=8, fig.width=8}



  bargraph_asv <- list()
  bargraph_species <- list()
  
  for (one_sample_type in sample_type_surface) {
    
      bargraph_asv [[one_sample_type]] <- phyloseq_long_bargraph(long[[one_sample_type]], text_scaling = 0.75,n_bars = 20, use_asv = TRUE, 
                                                                title = one_sample_type)
      bargraph_species[[one_sample_type]] <- phyloseq_long_bargraph(long[[one_sample_type]], text_scaling = 0.75,n_bars = 20 ,use_asv = FALSE, 
                                                                title = one_sample_type)

  }


```
## Barplots per sample

```{r fig.height=8, fig.width=8}

  bargraph_sample <- list()

  for (one_sample_type in sample_type_surface[4:11]) { 
    
    bargraph_sample[[one_sample_type]] <- plot_bar(ps[[one_sample_type]], x = "sample_label", fill = "class") + 
        geom_bar(aes(color = class, fill = class), stat = "identity", position = "stack") + 
        ggtitle(str_c("Class level - ", one_sample_type)) + theme(axis.text.y = element_text(size = 10)) + 
        theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) + coord_flip() + 
        scale_fill_viridis_d() + scale_color_viridis_d()
    print(bargraph_sample[[one_sample_type]])
}
```

## Heatmaps

Filter : Dictyo / Chrysophyceae /Pelago /Bacili / Crypto / Pyrami / Mamiello

### Class

#### Most abundant 10%

```{r fig.height=6, fig.width=9}
  heatmap_class_abundant <- list()

  for (one_sample_type in sample_type_surface[4:11]) { 
    
    ps_heat <- tax_glom(ps[[one_sample_type]],taxrank = "class") %>% 
      phyloseq_filter_abundant_taxa(fraction_min=0.1)
      
    heatmap_class_abundant[[one_sample_type]]  <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
                 taxa.label = "class",  taxa.order="division",
                 sample.label = "sample_label", sample.order="sample_label",
                 low="beige", high="red", na.value="beige", trans = scales::log_trans(10),
                 title=one_sample_type) + 
                 xlab("") + ylab("")
    
    # plot(heatmap(otu_table(ps_heat)))
    
    print(heatmap_class_abundant[[one_sample_type]] )
  
  }

```

#### Selected
```{r fig.height=6, fig.width=9}

class_selected <- c("Dictyochophyceae", "Chrysophyceae", "Pelagophyceae", "Bacillariophyta", "Cryptophyceae", "Pyramimonadales", "Mamiellophyceae")

  heatmap_class_selected <- list()

  for (one_sample_type in sample_type_surface[4:11]) { 
    
    reads_max <- case_when(str_detect(one_sample_type, "18S filter") ~ 30000,
                           str_detect(one_sample_type, "16S plastid") ~ 30000,
                           str_detect(one_sample_type, "18S sort") ~ 25000)
    
    ps_heat <- tax_glom(ps[[one_sample_type]],taxrank = "class") %>% 
      subset_taxa(class %in% class_selected )
    
    # Try to order by division and class...
    # tax_table <- data.frame(tax_table(ps_heat)@.Data)
    # taxa_names(ps_heat) <- str_c(tax_table$division, tax_table$class, sep=" - ") 
      
   gg  <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
                       taxa.label = "class",  taxa.order="class",
                       sample.label = "sample_label", sample.order="sample_label",
                       low="beige", high="red", na.value="gray95", trans=NULL, 
                       title=one_sample_type) + 
                       xlab("") + ylab("") +
                       theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust=1)) +
                       scale_fill_gradient(limits= c(0, reads_max), low="beige", high="red", na.value="gray95")
    
    if (str_detect(one_sample_type,"18S filter 0.2")) gg <- gg + geom_vline(xintercept = c(6.5, 9.5, 12.5))

    if (one_sample_type == "18S filter 3 um surface") gg <- gg +  geom_vline(xintercept = c(8.5, 10.5, 13.5))

    if (one_sample_type == "18S filter 20 um surface") gg <- gg +  geom_vline(xintercept = c(8.5, 9.5))

    
    # plot(heatmap(otu_table(ps_heat)))
    # , trans = scales::log_trans(10)
    
    print(gg)
    
    heatmap_class_selected[[one_sample_type]] <- gg
  
  }
```


### Species

#### Most abundant

```{r fig.height=6, fig.width=9}
  heatmap_genus_abundant <- list()

  for (one_sample_type in sample_type_surface[4:11]) { 
    
    ps_heat <- tax_glom(ps[[one_sample_type]],taxrank = "species") %>% 
      phyloseq_filter_abundant_taxa(fraction_min=0.1)
      
    heatmap_genus_abundant[[one_sample_type]]  <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
                 taxa.label = "species",  taxa.order="division",
                 sample.label = "sample_label", sample.order="sample_label",
                 low="beige", high="red", na.value="beige", 
                 title=one_sample_type)
    
    print(heatmap_genus_abundant[[one_sample_type]] )
  
  }
```

#### Selected

```{r fig.height=6, fig.width=9}

species_selected <- c("Thalassiosira_minima", "Fragilariopsis_cylindrus",  "Minidiscus_sp.", "Chaetoceros_neogracilis",
                      "Porosira_glacialis", "Coretrhon_inerme", "Palmaria_palmata", "Pseudo-nitzchia_seriata",
                      "Micromonas_polaris",  "Micromonas_clade_B3","Bathycoccus_prasinos", "Pyramimonas_gelidicola", 
                      "Geminigera_cryophila", "Pelagophyceae_XXX_sp." ,
                      "Phaeocystis_antarctica")


  heatmap_species_selected <- list()

  for (one_sample_type in sample_type_surface[4:11]) { 

reads_max <- case_when(str_detect(one_sample_type, "18S filter") ~ 20000,
                       str_detect(one_sample_type, "16S plastid") ~ 5000,
                       str_detect(one_sample_type, "18S sort") ~ 15000)    
    ps_heat <- tax_glom(ps[[one_sample_type]],taxrank = "species") %>% 
      subset_taxa(species %in% species_selected )
    
    # Try to order by division and species...
    # tax_table <- data.frame(tax_table(ps_heat)@.Data)
    # taxa_names(ps_heat) <- str_c(tax_table$division, tax_table$species, sep=" - ") 
      
    gg  <- plot_heatmap(ps_heat, method = "NMDS", distance = "bray", 
                       taxa.label = "species",  taxa.order="class",
                       sample.label = "sample_label", sample.order="sample_label",
                       low="beige", high="red", na.value="gray95", trans = NULL,
                       title=one_sample_type) + 
                       xlab("") + ylab("") +
                       theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust=1))+
                       scale_fill_gradient(limits= c(0,reads_max), low="beige", high="red", na.value="gray95")
      
    if (str_detect(one_sample_type, "18S filter 0.2")) gg <- gg + geom_vline(xintercept = c(6.5, 9.5, 12.5))

    if (one_sample_type == "18S filter 3 um surface") gg <- gg +  geom_vline(xintercept = c(8.5, 10.5, 13.5))

    if (one_sample_type == "18S filter 20 um surface") gg <- gg +  geom_vline(xintercept = c(8.5, 9.5))
    
                
    
    # plot(heatmap(otu_table(ps_heat)))
    # , trans = scales::log_trans(10)
    
    print(gg )
    
    heatmap_species_selected[[one_sample_type]] <- gg
  
  }
```

# NMDS


## Apply to all samples

```{r}
  nmds_sample <- list()

  for (one_sample_type in sample_type_surface[1:3]) { 
    
    one_ps <- ps[[one_sample_type]] %>% 
      tax_glom(taxrank = "species") 
    
    nmds_sample[[one_sample_type]] <- phyloseq_nmds(one_ps, 
                                                    title=one_sample_type, 
                                                    sample_color = season, 
                                                    sample_shape = fraction_name )
  }



```


# Vertical profile 

* Station 6 - 2015-01-16 
* Very important, must remove taxa that are not present in the filtered samples
* Do not use the TFF

## Filter the data
```{r}
  sample_type_profile <- str_c(sample_type_all, " profile")
  
  for (one_sample_type in sample_type_all) {
      
      one_sample_type_profile = str_c(one_sample_type, " profile")

      ps[[one_sample_type_profile]] <- ps[[one_sample_type]] %>% 
        subset_samples(station_id=="6") %>% 
        subset_samples(date =="2015-01-16") %>% 
        subset_samples(is.na(sample_concentration)) %>%
        filter_taxa(function(x) sum(x) > 0 , TRUE) %>% 
        phyloseq_normalize_median()
      
      # Rename the samples
      
      if (!str_detect(one_sample_type_profile, "sort")){
        sample_names(ps[[one_sample_type_profile]]) <- str_c(as.character(sample_data(ps[[one_sample_type_profile]])$depth), "m",
                                               as.character(sample_data(ps[[one_sample_type_profile]])$fraction_min),
                                               sep="_")
      } else {
        sample_names(ps[[one_sample_type_profile]]) <- str_c(as.character(sample_data(ps[[one_sample_type_profile]])$depth), "m",
                                       as.character(sample_data(ps[[one_sample_type_profile]])$fraction_name),
                                       sep="_")
      }
      
      
      print(glue("Phyloseq - {one_sample_type_profile}"))
      print(ps[[one_sample_type_profile]] )
      cat("============================\n")
      
  }

```

## Barplots per depth

```{r fig.height=8, fig.width=8}

  bargraph_sample <- list()

  for (one_sample_type in sample_type_profile[4:11]) { 
    
    bargraph_sample[[one_sample_type]] <- plot_bar(ps[[one_sample_type]], x = "depth", fill = "class") + 
        geom_bar(aes(color = class, fill = class), stat = "identity", position = "stack") + 
        ggtitle(str_c("Class level - ", one_sample_type)) + theme(axis.text.y = element_text(size = 10)) + 
        theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) + coord_flip() + 
        # scale_fill_viridis_d() + scale_color_viridis_d() + 
        scale_fill_manual(values = class_colors) +  scale_color_manual(values = class_colors) + 
        scale_x_reverse() + xlab("Depth (m)") + ylab("Reads") +
        theme_bw()
    
    print(bargraph_sample[[one_sample_type]])
}
```

# Compare the contributions of CLASS and SPECIS using the different methods

## Sample by sample comparison

* Each dot corresponds to the relative contribution of the taxonomic level considered for a sample for which both 18S and 16S has been performed

```{r fig.height=6, fig.width=10}
long_species_sample <- list()
long_class_sample <- list()

  for (one_sample_type in sample_type) {
      ps_one <- phyloseq_normalize_percent(ps[[one_sample_type]])
      
      long_one <- phyloseq_transform_to_long(ps_one) %>%
        group_by(date, station_id, depth, fraction_name, division, class, order, family, genus, species) %>%
        summarize(n_reads = sum(n_reads)) %>% 
        ungroup()
      
      long_species_sample[[one_sample_type]] <- long_one 
      
      long_one <- phyloseq_transform_to_long(ps_one) %>%
        group_by(date, station_id, depth, fraction_name, division, class) %>%
        summarize(n_reads = sum(n_reads)) %>% 
        ungroup()
      
      long_class_sample[[one_sample_type]] <- long_one 
    }
   
 long_compare_species_18S_16S <- full_join(rename(long_species_sample[["18S filter"]], n_reads_18S_filter = n_reads),
                                           rename(long_species_sample[["16S plastid"]],n_reads_16S_plastid = n_reads))
 
 ggplot(long_compare_species_18S_16S, aes(x = n_reads_18S_filter, y = n_reads_16S_plastid)) +
   geom_point(aes(color=class)) +
   scale_color_manual(values = class_colors) +
   ggtitle("Species")
 
 long_compare_class_18S_16S <- long_compare_species_18S_16S %>% 
        group_by(date, station_id, depth, fraction_name, division, class) %>%
        summarize(n_reads_18S_filter = sum(n_reads_18S_filter, na.rm=TRUE), n_reads_16S_plastid = sum( n_reads_16S_plastid, na.rm=TRUE)) 
 
  ggplot(long_compare_class_18S_16S, aes(x = n_reads_18S_filter, y = n_reads_16S_plastid)) +
   geom_point(aes(color=class)) +
   scale_color_manual(values = class_colors)+
   ggtitle("Classes")
   
```

## Define upset function
* Needs only to have 0 and 1 in the matrix

```{r}
plot_upset_long <- function(long_comparison, file_name){

long_upset <- long_comparison %>% 
  mutate_if(is.numeric, funs(case_when(is.na(.) ~ 0,
                                       TRUE ~ 1))) %>% 
  as.data.frame()

fig_upset <- UpSetR::upset(long_upset, empty.intersections = "on", number.angles = 30, point.size = 5, line.size = 2, 
    mainbar.y.label = "Species intersections", sets.x.label = "Species #", order.by = "freq",
    text.scale = c(2, 2, 2, 2, 2,2), mb.ratio = c(0.7, 0.3))

# text.scale = c(intersection size title, intersection size tick labels, set size title, set size tick labels, set names, numbers above bars)

print(fig_upset)
png(file_name, width = 1000, height = 600, units = "px")
print(fig_upset)
dev.off()
}
```


## Global comparison of species found with the 3 methods


* Compute table of number of samples for each species (rows) vs the three methods (columns)

* Only keep
  * 2015 samples because it is the only dataset for which we have the three types of samples
  * 0.2 and 3 um fractions (to be comparable with sorting)
* Remove species that contains _X


```{r  fig.height=6, fig.width=10}

long_species <- list()
long_class   <- list()

  for (one_sample_type in sample_type) {
      long_one <- long_species_sample[[one_sample_type]] %>%
        filter(date >= as.Date("2015-01-01")) %>% 
        filter(!str_detect(species, "_X")) %>% 
        filter(fraction_name != "> 20 um") %>% 
        group_by(division, class, order, family, genus, species) %>%
        summarize(n_reads = sum(n_reads), n_samples = n()) %>% 
        rename_at(vars(contains("n_")), funs(str_c( .,str_replace_all(one_sample_type, " ", "_"), sep="_"))) %>% 
        ungroup()
      
      long_species[[one_sample_type]] <- long_one 

      long_one <- long_class_sample[[one_sample_type]] %>%
        filter(date >= as.Date("2015-01-01"))%>% 
        filter(fraction_name != "> 20 um") %>%
        group_by(division, class) %>%
        summarize(n_reads = sum(n_reads), n_samples = n())%>% 
        rename_at(vars(contains ("n_")), funs( str_c(., str_replace_all(one_sample_type, " ", "_"), sep="_")))  %>% 
        ungroup()    
      
      long_class[[one_sample_type]] <- long_one 

  }

  long_class_comparison <- long_class %>% 
    purrr::reduce(full_join)
  long_species_comparison <- long_species %>% 
    purrr::reduce(full_join)
  
  cat("Class not found in one type of sample\n")
  cat("18S filter\n")
   kable(filter(long_class_comparison, is.na(n_samples_18S_filter)))
  cat("18S sort\n")
   kable(filter(long_class_comparison, is.na(n_samples_18S_sort)))
  cat("16S filter\n")
   kable(filter(long_class_comparison, is.na(n_samples_16S_plastid)))
   

  cat("Species  found in the three types of samples\n")
  cat("18S filter\n")
   kable(filter(long_species_comparison, (!is.na(n_samples_18S_filter)& 
                                        !is.na(n_samples_18S_sort) &
                                        !is.na(n_samples_16S_plastid) )))

  cat("Species only found in one type of sample\n")
  cat("18S filter\n")
   kable(filter(long_species_comparison, (is.na(n_samples_18S_sort) &
                                        is.na(n_samples_16S_plastid) )))
  cat("18S sort\n")
   kable(filter(long_species_comparison, (is.na(n_samples_18S_filter)&
                                        is.na(n_samples_16S_plastid) )))
  cat("16S filter\n")
   kable(filter(long_species_comparison, (is.na(n_samples_18S_filter)& 
                                        is.na(n_samples_18S_sort)  )))
 
   openxlsx::write.xlsx(list(long_class_comparison,long_species_comparison ), 
                        file="../dada2/method_comparison.xlsx",
                        sheetName=c("Class","Species"))
   
  long_species_comparison_upset <- long_species_comparison %>% 
  select(species, contains("samples")) %>%
  rename_at(vars(contains("n_")), funs(str_replace(.,"n_samples_", ""))) 
   
  plot_upset_long(long_species_comparison_upset, "fig/upset_method.png")

```


## Comparison of species found in the different fractions for 18S filter and 18S sorted   

* Compute table of number of samples for each species (rows) vs the three fractions (columns)

* Only keep
  * 2015 samples because it is the only dataset for which we have the three types of samples
  * Genera that do not contain _X


```{r  fig.height=6, fig.width=10}

long_species_fraction <- list()
long_class_fraction <- list()

  for (one_sample_type in sample_type[c(1,3)]) {
      long_one <- long_species_sample[[one_sample_type]] %>%
        filter(date >= as.Date("2015-01-01")) %>% 
        filter(!str_detect(species, "_X")) %>%  
        group_by(fraction_name, division, class, order, family, genus, species) %>%
        summarize(n_samples = n()) %>% 
        pivot_wider(names_from = fraction_name, values_from = n_samples) %>% 
        ungroup()
      
      long_species_fraction[[one_sample_type]] <- long_one 

      long_one <- long_class_sample[[one_sample_type]] %>%
        filter(date >= as.Date("2015-01-01")) %>% 
        group_by(fraction_name, division, class) %>%
        summarize(n_samples = n()) %>% 
        pivot_wider(names_from = fraction_name, values_from = n_samples) %>% 
        ungroup()
      
      long_class_fraction[[one_sample_type]] <- long_one 

  }

  long_class_fraction_comparison <- long_class_fraction %>% 
    purrr::reduce(full_join)
  long_species_fraction_comparison <- long_species_fraction %>% 
    purrr::reduce(full_join)
  
   kable(long_class_fraction_comparison)
   kable(long_species_fraction_comparison)
   
   plot_upset_long(long_species_fraction_comparison, "fig/upset_filter_sort.png")

```


## Comparison of species found in the different fractions for 18S filter   
* Compute table of number of samples for each species (rows) vs the three fractions (columns)

* Only keep
  * Genera that do not contain _X
* We consider now ALL samples (will need to remove samples for which the > 20 um is missing)

```{r  fig.height=6, fig.width=10}

long_species_fraction <- list()
long_class_fraction <- list()

  for (one_sample_type in sample_type[c(1)]) {
      long_one <- long_species_sample[[one_sample_type]] %>%
        filter(!str_detect(species, "_X")) %>%  
        group_by(fraction_name, division, class, order, family, genus, species) %>%
        summarize(n_samples = n()) %>% 
        pivot_wider(names_from = fraction_name, values_from = n_samples) %>% 
        ungroup()
      
      long_species_fraction[[one_sample_type]] <- long_one 

      long_one <- long_class_sample[[one_sample_type]] %>%
        group_by(fraction_name, division, class) %>%
        summarize(n_samples = n()) %>% 
        pivot_wider(names_from = fraction_name, values_from = n_samples) %>% 
        ungroup()
      
      long_class_fraction[[one_sample_type]] <- long_one 

  }

  long_class_fraction_comparison <- long_class_fraction[[one_sample_type]]

  long_species_fraction_comparison <- long_species_fraction[[one_sample_type]] 

   kable(long_class_fraction_comparison)
   kable(long_species_fraction_comparison)
   
   plot_upset_long(long_species_fraction_comparison, "fig/upset_filter.png")
```


## Comparison of species found in the different fractions for 18S sort   
* Compute table of number of samples for each species (rows) vs the three fractions (columns)

* Only keep
  * Genera that do not contain _X
* We consider now ALL samples (will need to remove samples for which the > 20 um is missing)

```{r  fig.height=6, fig.width=10}

long_species_fraction <- list()
long_class_fraction <- list()

  for (one_sample_type in sample_type[c(3)]) {
      long_one <- long_species_sample[[one_sample_type]] %>%
        filter(!str_detect(species, "_X")) %>%  
        group_by(fraction_name, division, class, order, family, genus, species) %>%
        summarize(n_samples = n()) %>% 
        pivot_wider(names_from = fraction_name, values_from = n_samples) %>% 
        ungroup()
      
      long_species_fraction[[one_sample_type]] <- long_one 

      long_one <- long_class_sample[[one_sample_type]] %>%
        group_by(fraction_name, division, class) %>%
        summarize(n_samples = n()) %>% 
        pivot_wider(names_from = fraction_name, values_from = n_samples) %>% 
        ungroup()
      
      long_class_fraction[[one_sample_type]] <- long_one 

  }

  long_class_fraction_comparison <- long_class_fraction[[one_sample_type]]

  long_species_fraction_comparison <- long_species_fraction[[one_sample_type]] 

   kable(long_class_fraction_comparison)
   kable(long_species_fraction_comparison)
   
   plot_upset_long(long_species_fraction_comparison, "fig/upset_sort.png")
```


# Figures

## Figure 2 - Treemap

```{r fig.height=30, fig.width=15}

treemap_18S_filt <- cowplot::plot_grid(treemap_class[["18S filter 0.2 um surface"]]$gg,
                                       treemap_class[["18S filter 3 um surface"]]$gg,
                                       treemap_class[["18S filter 20 um surface"]]$gg, 
                                       ncol=1)

treemap_18S_sort <- cowplot::plot_grid(treemap_class[["18S sort pico surface"]]$gg,
                                       treemap_class[["18S sort nano surface"]]$gg,
                                       treemap_class[["fill in"]]$gg,
                                       ncol=1)

treemap_16S_filt <- cowplot::plot_grid(treemap_class[["16S plastid 0.2 um surface"]]$gg,
                                       treemap_class[["16S plastid 3 um surface"]]$gg,
                                       treemap_class[["16S plastid 20 um surface"]]$gg, 
                                       ncol=1)

fig_2 <- cowplot::plot_grid(bargraph_species[["18S filter surface"]]$gg, treemap_18S_filt,
                            bargraph_species[["18S sort surface"]]$gg, treemap_18S_sort,
                            bargraph_species[["16S plastid surface"]]$gg, treemap_16S_filt,
                            ncol=2,rel_widths = c(2,1))

fig_2
```

## Figure Supp - Treemap at genus level

```{r fig.height=15, fig.width=15}

treemap_18S_filt <- cowplot::plot_grid(treemap_genus[["18S filter 0.2 um surface"]]$gg,
                                       treemap_genus[["18S filter 3 um surface"]]$gg,
                                       treemap_genus[["18S filter 20 um surface"]]$gg, 
                                       ncol=1)

treemap_18S_sort <- cowplot::plot_grid(treemap_genus[["18S sort pico surface"]]$gg,
                                       treemap_genus[["18S sort nano surface"]]$gg,
                                       treemap_genus[["fill in"]]$gg,
                                       ncol=1)

treemap_16S_filt <- cowplot::plot_grid(treemap_genus[["16S plastid 0.2 um surface"]]$gg,
                                       treemap_genus[["16S plastid 3 um surface"]]$gg,
                                       treemap_genus[["16S plastid 20 um surface"]]$gg, 
                                       ncol=1)

fig_supp_treemap <- cowplot::plot_grid(treemap_18S_filt,
                            treemap_18S_sort,
                            treemap_16S_filt,
                           labels = c("A" ,"B", "C"), label_x = 0.9,
                            ncol=3)

fig_supp_treemap
```


## Figure 3 - NMDS 18S filter

```{r fig.height=12, fig.width=8}


fig_3 <- cowplot::plot_grid(nmds_sample[["18S filter surface"]]$gg_samples,
                            nmds_sample[["18S filter surface"]]$gg_taxa, 
                           labels = c("A" ,"B"), label_x = 0.9,
                            nrow=2,rel_widths = c(1,1))

fig_3
```

## Figure sup - NMDS 16S plastid

```{r fig.height=12, fig.width=8}


fig_3 <- cowplot::plot_grid(nmds_sample[["16S plastid surface"]]$gg_samples,
                            nmds_sample[["16S plastid surface"]]$gg_taxa, 
                           labels = c("A" ,"B"), label_x = 0.9,
                            nrow=2,rel_widths = c(1,1))

fig_3
```

## Figure 4 - Heatmap 18S filter

```{r fig.height=12, fig.width=25}
fig_4 <- cowplot::plot_grid(heatmap_class_selected[["18S filter 0.2 um surface"]],
                           heatmap_class_selected[["18S filter 3 um surface"]],
                           heatmap_class_selected[["18S filter 20 um surface"]],
                           heatmap_species_selected[["18S filter 0.2 um surface"]],
                           heatmap_species_selected[["18S filter 3 um surface"]],
                           heatmap_species_selected[["18S filter 20 um surface"]], 
                           labels = c("A" ,"B","C", "D","E", "F"), label_x = 0.9,
                           nrow=2, ncol=3)
fig_4
```


## Figure sup - Heatmap 18S sort

```{r fig.height=10, fig.width=15}
fig_heatmap_sort <- cowplot::plot_grid(heatmap_class_selected[["18S sort pico surface"]],
                           heatmap_class_selected[["18S sort nano surface"]],
                           heatmap_species_selected[["18S sort pico surface"]],
                           heatmap_species_selected[["18S sort nano surface"]], 
                           labels = c("A" ,"B","C", "D"), label_x = 0.9,
                           nrow=2, ncol=2)
fig_heatmap_sort
```

## Figure 6 - Vertical Profile

```{r fig.height=12, fig.width=6}

legend <- cowplot::get_legend( bargraph_sample[["18S filter 0.2 um profile"]] + 
                      # create some space to the left of the legend    
                      theme(legend.box.margin = margin(0, 0, 0, 20))
                    )
fig_6 <- cowplot::plot_grid(bargraph_sample[["18S filter 0.2 um profile"]]+ theme(legend.position="none"),legend,
                            bargraph_sample[["18S filter 3 um profile"]] + theme(legend.position="none"),NULL,
                            bargraph_sample[["18S filter 20 um profile"]]+ theme(legend.position="none"),NULL,
                            labels = c("" ,"A","", "B","", "C"), label_x = 0.9,
                            nrow=3, ncol=2, rel_widths = c(3, 1.7))
fig_6
```

## Figure X - Upset R

* Problem is that upset does not provide a ggplot object.  So need to save as png
* Use ggplotify library : https://cran.r-project.org/web/packages/ggplotify/vignettes/ggplotify.html
* But it heats a bit the plots

```{r fig.height=10, fig.width=6}

p1 <- cowplot::ggdraw() + cowplot::draw_image("fig/upset_method.png")
p2 <- cowplot::ggdraw() + cowplot::draw_image("fig/upset_filter.png")
p3 <- cowplot::ggdraw() + cowplot::draw_image("fig/upset_sort.png")

fig_X_upset <- cowplot::plot_grid(p1, p2, p3, labels = c("A", "B", "C"),label_x = 0.9,
                                  nrow=3, ncol=1)
fig_X_upset
```

# CARBOM code

```{r, eval = FALSE, echo=FALSE}
# === 
# Convert from phyloseq to vegan
# ===

# convert the sample_data() within a phyloseq object to a vegan compatible data object
pssd2veg <- function(physeq) {
  sd <- sample_data(physeq)
  return(as(sd,"data.frame"))
}

# convert the otu_table() within a phyloseq object to a vegan compatible data object
psotu2veg <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(as(OTU, "matrix"))
}

# Create also a function to rename the samples and the taxa...


# =======================================================
# 18S Compute matrixes of abundances for OTUs for autotrophs
#-----------------------------------------------------------

# Reshape the abundance matrix in condensed form 
  # Keep only necessary columns
  otu_matrix <- otu_melt_auto %>% transmute(sample_label, otu_label, sequence_number, Final_Taxo2, Final_Taxo3, Final_Taxo4, Final_Taxo7, Final_Taxo8)
  # Reshape in wide form, sort by taxonomy and then remove the taxonomy information
  otu_matrix <- otu_matrix %>% 
                spread(sample_label, sequence_number) %>% 
                arrange(Final_Taxo2, Final_Taxo3, Final_Taxo4, otu_label ) %>% 
                select (-Final_Taxo2, -Final_Taxo3, -Final_Taxo4, -Final_Taxo7, -Final_Taxo8)
  
# Provide row.names using the labels for the samples  
  row.names(otu_matrix)=otu_matrix$otu_label
  
# Remove the column that contains the labels
  otu_matrix=select(otu_matrix, -otu_label)  

# Transpose the matrix
  otu_matrix=t(otu_matrix)  

    
# Normalize the data per sample
  otu_matrix_relative <- otu_matrix/rowSums(otu_matrix)
  
# determine the maximum relative abundance for each column
  maxab <- apply(otu_matrix_relative, 2, max)  # 1 indicates row and 2 columns
  
# remove the otus with less than 20 % as their maximum relative abundance -> for heatmaps
  otu_low <- names(which(maxab < 0.20))
  otu_abund <- names(which(maxab >= 0.20))
  otu_matrix_relative_abund <- otu_matrix_relative[, -which(colnames(otu_matrix_relative) %in% otu_low)]
  otu_matrix_abund <- otu_matrix[, -which(colnames(otu_matrix_relative) %in% otu_low)]

# Export matrices
  write.table(otu_matrix_relative, file="otu_matrix_relative.txt",
              row.names = TRUE,  col.names = TRUE, sep="\t", quote=FALSE)
  write.table(otu_matrix_relative_abund, file="otu_matrix_relative_abund.txt",
              row.names = TRUE,  col.names = TRUE, sep="\t", quote=FALSE)

# Do a small table with the information from major OTUs
  otu_table_paper <- otu_table %>% filter(otu_label %in% otu_abund) %>% transmute(otu, otu_total, Division=Final_Taxo3, Class=Final_Taxo4, Genus=Final_Taxo7, Species=Final_Taxo8, BLAST_best_Accession, BLAST_best_identity, BLAST_best_Description) %>% arrange(otu)
  write.table(otu_table_paper, file="otu_18S_table_paper.txt", sep = "\t", row.names = FALSE, quote=FALSE)

# Number of OTU per class
  otu_table_class <- otu_table %>% group_by(Final_Taxo3, Final_Taxo4) %>% summarize(number_of_otu = n(), number_of_reads = sum(otu_total)) 
  
# =======================================================
# 18S  - Similarity Matrix for samples  : Compute it for different metrix 
#-----------------------------------------------------------
  
# Compute dissimil between samples using Bray-Curtis on non normalized matrix (can lead to problems - see paper in Nature Microbio reviews)
  sample_dissimil_bray=vegdist(otu_matrix_abund,method = "bray", na.rm = F)
  
# Compute dissimil between otus using Bray-Curtis
  otu_dissimil_bray=vegdist(t(otu_matrix_abund),method = "bray", na.rm = F)
  
# Change from dissimil to simil and transform to matrix
  otu_simil_bray <- as.matrix(1 - otu_dissimil_bray)

# =======================================================
# 18S - Compute Shannon diversity based on otu 
#-----------------------------------------------------------

  otu_diversity<-as.data.frame(vegan::diversity(otu_matrix, "shannon"))

# =======================================================
#  18S - Fig. 4 - Heat maps based on tutorial  : #  http://www.molecularecologist.com/2013/08/making-heatmaps-with-r-for-microbiome-analysis/  
#-----------------------------------------------------------

  
# Do average linkage hierarchical clustering of the rows. Other options are 'aver', 'complete' or 'single'
  sample_cluster=hclust(sample_dissimil_bray, method = "complete")
  plot(sample_cluster, cex=1)

# Do average linkage hierarchical clustering of the otus. Other options are 'complete' or 'single'
  otu_cluster <- hclust(otu_dissimil_bray, "complete")
  plot(otu_cluster, cex=1)

# Do the heatmap with the samples and otu clustered
  heatmap(as.matrix(otu_matrix_relative_abund), Rowv = as.dendrogram(sample_cluster), Colv = as.dendrogram(otu_cluster), col = scaleyellowred, margins = c(12, 4))
  
  
# Do the heatmap with the samples only clustered
  heatmap(as.matrix(otu_matrix_relative_abund), Rowv = as.dendrogram(sample_cluster), Colv = NA, col = scaleyellowred, margins = c(20, 4))


# Fig. 4 - Do the heatmap with the samples only clustered Better to provide the scale
  pdf("./pdf_figures_2.0/Fig_4 Heatmap 18S  2.0.pdf",width = 8, height = 10, useDingbats=FALSE)
  heatmap.2(as.matrix(otu_matrix_relative_abund), 
            Rowv = as.dendrogram(sample_cluster), Colv = FALSE, 
            dendrogram="row", col = scaleyellowred, trace="none", 
            key = TRUE, density.info="none", cexRow=0.8, cexCol=0.8, margins = c(15, 10), 
            lmat = rbind(c(0,3),c(2,1),c(0,4)), lwid = c(1,4), lhei = c(0.2,4,0.6), srtCol=45, 
            key.title= NA , key.xlab = "Relative abundance")
  dev.off()
  

# 18S - Stats - MSDS based using vegan on Bray distance of relative abundance --------


# Create environmental table to put on msds

  envdata_table <- select(filter(samples,Select_18S_nifH == "Yes"), sample_illumina, depth, 
                          transect_distance, phosphates, silicates, nitrates, temperature, 
                          fluorescence, salinity, fraction, 
                          bacteria, prochlorococcus, synechococcus, picoeukaryotes, nanoeukaryotes)
  envdata_table <- envdata_table %>%  mutate (N_P =nitrates/phosphates) %>% 
                                      select(-bacteria, -fraction, -nitrates, -phosphates, - silicates)
  
  row.names(envdata_table) <- envdata_table$sample_illumina
  envdata_table <- select(envdata_table, -sample_illumina)
    # write.table(envdata_table, file="envdata_table.txt", 
    #           row.names = TRUE,  col.names = TRUE, sep="\t", quote=FALSE)
  # envdata_table <- envdata_table[ order(row.names(envdata_table)), ]  # Note if reorder the rows then the row name disappear...
  
  # Rename the row names to be shorter
  otu_matrix_relative_abund_msds <- otu_matrix_relative_abund
  row_names_short <- str_split_fixed(row.names(otu_matrix_relative_abund_msds),"_", n=4)
  row.names(otu_matrix_relative_abund_msds) <- row_names_short[,4]
  otu_matrix_relative_abund_msds <- otu_matrix_relative_abund_msds[ order(row.names(otu_matrix_relative_abund_msds)), ]
  
# Compare the different distances for separation - Bray Curtis appears best
  rankindex(envdata_table, otu_matrix_relative_abund_msds)
  
#         euc         man         gow         bra         kul 
# -0.10460145 -0.12017731 -0.15173360 -0.04435000 -0.04604831 
  
# Compute MSDS
  otu_matrix_relative_abund.mds <- metaMDS(otu_matrix_relative_abund_msds, trace = TRUE) 
  otu_matrix_relative_abund.mds
  # plot(otu_matrix_relative_abund.mds, type = "t", display="species")
  plot(otu_matrix_relative_abund.mds, type = "t", display="sites")
  # plot(otu_matrix_relative_abund.mds, type = "p", display="sites")

# Fit environmental variables including Pico and Nano
  otu_matrix_relative_abund.ef <- envfit(otu_matrix_relative_abund.mds, envdata_table, permu = 999, na.rm = TRUE)
  otu_matrix_relative_abund.ef
  plot(otu_matrix_relative_abund.ef, p.max = 1)
  
# global Multidimensional Scaling using monoMDS
# 
# Data:     otu_matrix_relative_abund_msds 
# Distance: bray 
# 
# Dimensions: 2 
# Stress:     0.2500929 
# Stress type 1, weak ties
# No convergent solutions - best solution after 20 tries
# Scaling: centring, PC rotation, halfchange scaling 
# Species: expanded scores based on 'otu_matrix_relative_abund_msds'***VECTORS

#                      NMDS1    NMDS2     r2 Pr(>r)  
# depth             -0.99531  0.09678 0.0578  0.370  
# transect_distance  0.58527 -0.81084 0.0890  0.192  
# temperature        0.99968 -0.02541 0.0476  0.459  
# fluorescence      -0.97532  0.22081 0.2018  0.017 *
# salinity           0.13213 -0.99123 0.0430  0.475  
# prochlorococcus    0.09546 -0.99543 0.0458  0.441  
# synechococcus     -0.62789  0.77830 0.0487  0.409  
# picoeukaryotes    -0.96605 -0.25837 0.0098  0.838  
# nanoeukaryotes    -0.62715  0.77890 0.0788  0.243  
# N_P               -0.65297 -0.75738 0.0346  0.524  
# ---
# Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
# Permutation: free
# Number of permutations: 999 
  

# PCA (principal compoment) analysis -----------------------------------------------------

otu_matrix_relative_abund.rda <- rda(otu_matrix_relative_abund_msds, scale=TRUE)
otu_matrix_relative_abund.rda
plot(otu_matrix_relative_abund.rda)

# CCA (correspondencen) analysis -----------------------------------------------------

otu_matrix_relative_abund.cca <- cca(otu_matrix_relative_abund_msds, scale=TRUE)
otu_matrix_relative_abund.cca
plot(otu_matrix_relative_abund.cca)

# Mutiple correlation -----------------------------------------------------
otu_matrix_relative_abund_msds.cor <- corr.test(envdata_table, otu_matrix_relative_abund_msds, use="na.or.complete", method="spearman")
write.table(otu_matrix_relative_abund_msds.cor$r, file="carbom_18S_mutiple_correlation.txt", sep="\t", quote = FALSE)
write.table(otu_matrix_relative_abund_msds.cor$p, file="carbom_18S_mutiple_correlation_pvalue.txt", sep="\t", quote = FALSE)
```

